---
title: "R Notebook Sandbox"
output: 
  html_notebook:
    df_print: paged
---

```{r}
#library(devtools)
#local.path = "C:/Users/Dorbs of Doom/_git_/WSU_STATS419_FALL2020/"
#source_url(paste0(github.path, "master/functions/libraries.R"), local=T)
#source_url(paste0(github.path, "master/functions/functions-imdb.R"), local=T)

#source("functions/libraries.R", local=F)
#source("functions/functions-imdb.R")
```

##02_datasets
#the "rotate matrix" functions

```{r matrix}
myMatrix=matrix(c(1,0,2,
                  0,3,0,
                  4,0,5), nrow=3, byrow=T);


transposeMatrix=function(mat)
{ 
  t(mat);
}

rotateMatrix90=function(mat)
{
  tMatrix=matrix(c(0,0,1,
                   0,1,0,
                   1,0,0), nrow=3, byrow=T)
  x=tMatrix%*%mat
  transposeMatrix(x)
}

rotateMatrix90(myMatrix)

rotateMatrix180=function(mat)
{
  ninety=rotateMatrix90(mat)
  tMatrix=matrix(c(0,0,1,
                   0,1,0,
                   1,0,0), nrow=3, byrow=T)
  x=tMatrix%*%ninety
  transposeMatrix(x)
}

rotateMatrix180(myMatrix)

rotateMatrix270=function(mat)
{
  oneEighty=rotateMatrix180(mat)
  tMatrix=matrix(c(0,0,1,
                   0,1,0,
                   1,0,0), nrow=3, byrow=T)
  x=tMatrix%*%oneEighty
  transposeMatrix(x)
}

rotateMatrix270(myMatrix)
```

#the iris visualization reproduction

```{r iris}
#-----------------------------------------
iris
class(iris)

my_cols <- c("red", "lime green", "blue")  

pairs(~Sepal.Length+Sepal.Width+Petal.Length+Petal.Width, data=iris, 
      main="Iris Data (red=setosa,green=versicolor,blue=virginica",
      pch=21, bg=my_cols[iris$Species], col="black", 
      cex=1.4, cex.labels=2, cex.main=1.75, cex.axis=1.75 )
```

Notes:
pch changes the shape and filling of plot points
bg background color of the plot points
col border color of the plot points
cex size of the plot points
cex.labels size of the variables (sepal.length, sepal.width, petal.length, petal.width)
cex.main size of title
font.labels changes the font
cex.axis changes size of axis font
typeof() tells you what type the variable is

#clean Personality dataset

```{r personality}
install.packages("readr", dependencies=TRUE)  #need to to use read_delim
install.packages("dplyr", dependencies = TRUE)
library(readr)
library(dplyr)

#read txt file, remove V00 
personality_dat<-read_delim("personality-raw.txt", delim='|')
remove_V00_dat<-subset(personality_dat, select = -c(V00))

#remove date_test column to turn into year and week columns
date_col<-remove_V00_dat[[2]]

just_date<-gsub(" .*", "", date_col); just_date
formatted_date<-as.Date(just_date, format="%m/%d/%Y"); formatted_date

year_col<-format(formatted_date, format="%Y")
week_col<-format(formatted_date, format="%V")

#remove date_test column from remove_V00_dat and create a frame
remove_date_col<-subset(remove_V00_dat, select=-c(date_test))
asframe<-as.data.frame(remove_date_col, stringsAsFactors=FALSE)

#add year and week columns to remove_dat_col
added_year<-cbind(asframe, Year=year_col)
added_week<-cbind(added_year, Week=week_col)
reorder<- added_week[, colnames(added_week)[c(1, 62, 63, 2:61 )]]

#descending order by date
descending<-reorder[order(-xtfrm(reorder[,2]), -xtfrm(reorder[,3])), ]


#filter for duplicates and finalize cleaned data
personality_clean<- descending %>% distinct(md5_email, .keep_all=TRUE)

#export to txt file
write.table(personality_clean, file="personality_clean.txt", 
            sep="|", dec=".", row.names=FALSE, col.names=TRUE)

```

#Personality dataset statistics

```{r personality statistics}
#using personality_clean

#extract monte.schaffer@gmail.com row (1st row)
monte<-personality_clean %>% slice(1:1)

doSampleVariance = function(x, method)
{
  if(method=="naive")
  {
    count<-0
    sum<-0
    sumsq<-0
    for (val in x) {
      count<-count+1
      sum<-sum+val
      sumsq<-sumsq+val*val
    }
    var<-(sumsq-(sum*sum)/count)/(count-1) 
    summary<-c("sum"=sum, "sumSq"=sumsq,
               "var"=var)
    summary
    
  }
  else
  {
    # two-pass algorithm
    n<-0
    sum1<-0
    sum2<-0
    for (val in v) {
      n<- n+1
      sum1<-sum1+val
    }
    
    mean<-sum1/n
    
    for (val in v) {
      sum2<-sum2+((val-mean)^2)
    }
    var<-sum2 / (n-1)
    
    result<-c("sum"=sum1, "sum2"=sum2, "var"=var)  
    result
  }
}

doMode = function(x)
{
  # This function finds highest frequencies
  # using frequency table() and max()
  
  t<-table(x)
  xmax<-max(t)
  t; xmax
  
  if (all(t == xmax))
  {
    mode = NA
  } else if (is.numeric(x))
  {
    mode = as.numeric(names(t)[t == xmax])
  } else
  {
    mode = names(t)[t == xmax]
  }
  mode
}


doSummary = function(x)
{
  v<- as.double(subset(x, select=-c(md5_email, Year, Week)))
  # length
  length = length(v); length
  # number of NAs
  numNAs =sum(is.na(x$row)); numNAs
  # mean
  mean = mean(v)
  # median
  median=median(v)
  # mode 
  mode=doMode(v)
  # variance 
  var = doSampleVariance(v, "naive")
  # sd ... built in function but compare it to the custom function ... 
  r_sd<-sd(v)
  custom_sd<-sqrt(var[3])
  
  #summary
  summary<-c("length"=length, "number of NAs"=numNAs, 
                "mean"=mean, "median"=median, "mode"=mode,
                "var"=var, "built-in sd"=r_sd, "custom sd"=custom_sd)

  summary
}

doSummary(monte)

#table() will do a frequency table of the variable and its frequency
#can also use it for cross table( frequency with proportion), 2 way cross table
#or two way frequency table along with proportion in R
#even 3 way frequncy or cross tables

#--------------------------------------------------------------------
#Create Z-scores for monte
#plot z-scores (y) vs row scores (x)

x<- as.double(subset(monte, select=-c(md5_email, Year, Week)))
doZscore=function(x) 
{
  mean<-mean(x)
  sd<-sd(x)
  z<-(x-mean)/sd
  z
}

y<-doZscore(x)
plot(x,y)
lm(y~x)
```

#imdb example

```{r imdb}


## functions

education = function(one)
{
  result = list();
  result$who 		= one;
  result$think 	= c("intensitively", "critically");
  result$goal 	= "intelligences + character";
  result;	
}


me = education("monte");

# n > 1 ... 

# Research question:  who is a better actor?  Will Smith?  Denzel Washington?

##
myMatrix = matrix ( c (
  1, 0, 2,
  0, 3, 0,
  4, 0, 5
), nrow=3, byrow=T);

transposeMatrix = function(mat)
{
  t(mat);	
}

#rotateMatrix90(mat)	
#rotateMatrix180(mat)	
#rotateMatrix270(mat)
# 3x3 matrix ... ## matrix multiplication

# install.packages("stringr", dependencies=T);
library(stringr);	
# install.packages("rvest", dependencies=T);
library(rvest);	

# Research question:  who is a better actor?  Will Smith?  Denzel Washington?

## actor ... person_id
##			movie_id ... details
##			name ... count movies

# https://rvest.tidyverse.org/index.html

## Denzel Washington [nm0000243] vs Will Smith [nm0000226]

## https://www.imdb.com/filmosearch/?explore=title_type&role=nm0000243&ref_=filmo_ref_typ&sort=num_votes,desc&mode=detail&page=1&title_type=movie

# R , javascript, php, (c/c++)

# imdb ... 


grabFilmInfoFromFilmsPage = function(page)
{
  # 50 elements
  # # title = id = rank = year = rating = minutes = genre = votes = metascore = desc = millions
  
  movies = page %>%
    html_nodes(".mode-detail");
  
  
  
  pagecount = length(movies);
  
  result = data.frame( 			matrix(ncol = 11,nrow = pagecount) );
  # a matrix-type form with lots of NA values ...
  
  colnames(result) = c("rank", "title", "ttid", "year", "rated", "minutes", "genre", "ratings", "metacritic", "votes", "millions"); 
  
  
  for(i in 1:pagecount)
  {
    movie = movies[i];
    
    rank = movie %>%
      html_node(".lister-item-index") %>%
      html_text() %>%
      as.numeric();
    result$rank[i] = rank;
    
    title = movie %>%
      html_node(".lister-item-header a") %>%
      html_text();
    result$title[i] = title;
    
    ttid = movie %>%
      html_node(".lister-item-header a") %>%
      html_attr("href");
    
    temp = strsplit(ttid,"/",fixed=T);
    ttid = temp[[1]][3];
    result$ttid[i] = ttid;
    
    year = movie %>%
      html_node(".lister-item-year") %>%
      html_text();
    year = cleanupYear(year);
    result$year[i] = year;
    
    rated = movie %>%
      html_node(".certificate") %>%
      html_text();
    result$rated[i] = rated;
    
    minutes = movie %>%
      html_node(".runtime") %>%
      html_text();
    minutes = cleanupMinutes(minutes);
    result$minutes[i] = minutes;		
    
    genre = movie %>%
      html_node(".genre") %>%
      html_text();
    genre = str_trim(genre);
    result$genre[i] = genre;
    
    ratings = movie %>%
      html_node("div .rating-list") %>%
      html_attr("title");
    temp = strsplit(ratings,"/",fixed=T);
    temp = gsub("Users rated this","",temp[[1]][1],fixed=T);	
    temp = str_trim(temp);
    ratings = as.numeric(temp);
    result$ratings[i] = ratings;
    
    metacritic = movie %>%
      html_node(".ratings-metascore span") %>%
      html_text();
    metacritic = as.numeric(str_trim(metacritic));
    result$metacritic[i] = metacritic;
    
    # para ... +5 EASTER EGG ...
    
    info = movie %>%
      html_nodes(".lister-item-content p span") %>%
      html_text();
    
    votes = as.numeric(gsub(",","",info[8],fixed=T));
    result$votes[i] = votes;
    
    millions = cleanupMillions(info[11]);
    result$millions[i] = millions;			
  }
  
  #str(result
  
  
  result;
}







cleanupMillions = function(millions)
{
  millions = gsub('$','',millions, fixed=T);
  millions = gsub('M','',millions, fixed=T);
  
  millions = as.numeric(millions);
  millions;
}

cleanupMinutes = function(minutes)
{
  minutes = gsub('min','',minutes, fixed=T);
  
  minutes = as.numeric(minutes);
  minutes;
}

cleanupYear = function(year)
{
  year = gsub('(','',year, fixed=T);
  year = gsub(')','',year, fixed=T);
  year = gsub('I','',year, fixed=T);
  year = as.numeric(year);
  year;
}

grabNameFromFilmsPage = function(page)
{
  name = page %>%
    html_node(".header") %>%
    html_text();
  
  name = gsub("Most Rated Feature Films With","",name,fixed=T);
  name = str_trim(name);
  
  name;
}


grabFilmCountFromFilmsPage = function(page)
{
  totalcount = page %>%
    html_nodes(".desc") %>%
    html_text();
  
  temp = strsplit(totalcount,"of",fixed=T);
  temp2 = strsplit(temp[[1]][2],"titles", fixed=T);
  
  totalcount = str_trim(temp2[[1]][1]);
  totalcount = as.numeric(totalcount);
  
  temp2 = strsplit(temp[[1]][1],"to", fixed=T);
  
  pagecount = str_trim(temp2[[1]][2]);
  pagecount = as.numeric(pagecount);
  
  result = list();
  
  result$totalcount = totalcount;
  result$pagecount = pagecount;
  
  result;
}


#   nmid = "nm0000226";
# 	will = grabFilmsForPerson(nmid);
# 	plot(will$movies.50[,c(1,6,7:10)]);
#  	boxplot(will$movies.50$millions);

#   nmid = "nm0000243";
# 	denzel = grabFilmsForPerson(nmid);
# 	plot(denzel$movies.50[,c(1,6,7:10)]);
#  	boxplot(denzel$movies.50$millions);


# https://www.imdb.com/title/tt0466839/?ref_=filmo_li_tt ... get box office budget/gross if NA ... on millions. ..

grabFilmsForPerson = function(nmid)
{
  url = paste("https://www.imdb.com/filmosearch/?explore=title_type&role=",nmid,"&ref_=filmo_ref_typ&sort=num_votes,desc&mode=detail&page=1&title_type=movie", sep="");
  
  page1 = read_html(url);
  result = list();
  ## useful for other data purposes
  result$nmid = nmid;
  
  ## name of person
  result$name = grabNameFromFilmsPage(page1);
  result$countfilms = grabFilmCountFromFilmsPage(page1);
  
  result$movies.50 = grabFilmInfoFromFilmsPage(page1);
  
  
  
  
  ##  parallel format ...
  # ranks = page1 %>%
  # html_nodes(".lister-item-index") %>%
  # html_text() %>%
  # as.numeric();	
  
  # ranks;
  
  # years = page1 %>%
  # html_nodes(".lister-item-year") %>%
  # html_text();
  
  # years = gsub('(','',years, fixed=T);
  # years = gsub(')','',years, fixed=T);
  # years = gsub('I','',years, fixed=T);
  # years = as.numeric(years);
  
  # titles = page1 %>%	
  # html_nodes(".lister-item-header a") %>%
  # html_text();
  
  # titles;
  
  
  result;
}

#box plots for will smith
nmid = "nm0000226";
will = grabFilmsForPerson(nmid);
plot(will$movies.50[,c(1,6,7:10)]);
boxplot(will$movies.50$millions);
widx =  which.max(will$movies.50$millions);
will$movies.50[widx,];
summary(will$movies.50$year);  # bad boys for life ... did data change?

#box plot for denzel
nmid = "nm0000243";
denzel = grabFilmsForPerson(nmid);
plot(denzel$movies.50[,c(1,6,7:10)]);
boxplot(denzel$movies.50$millions);
didx =  which.max(denzel$movies.50$millions);
denzel$movies.50[didx,];
summary(denzel$movies.50$year);

#side by side box plots for Will and Denzel
par(mfrow=c(1,2));
boxplot(will$movies.50$millions, main=will$name, ylim=c(0,360), ylab="Raw Millions" );
boxplot(denzel$movies.50$millions, main=denzel$name, ylim=c(0,360), ylab="Raw Millions" );

#side by side box plots with adjust for inflation 
par(mfrow=c(1,2));
boxplot(will.2$dollar_adjusted, main=will$name, ylim=c(0,500), ylab="Adjusted Millions to 2020")
boxplot(denzel.2$dollar_adjusted, main=denzel$name, ylim=c(0,360), ylab="Adjusted Millions to 2020")

#side by side box plots for ratings
par(mfrow=c(1,2));
boxplot(will$movies.50$ratings, main=will$name, ylim=c(1,10), ylab="Ratings 1-10")
boxplot(denzel$movies.50$ratings, main=denzel$name, ylim=c(1,10), ylab="Ratings 1-10")

#side by side box plots for minutes
par(mfrow=c(1,2));
boxplot(will$movies.50$minutes, main=will$name, ylim=c(50,200), ylab="Movie Length")
boxplot(denzel$movies.50$minutes, main=denzel$name, ylim=c(50,202), ylab="Movie Length")


------------------------------------------------
  
# https://www.in2013dollars.com/us/inflation/2000?endYear=1982&amount=100
# create variable $millions.2000 to convert all money to  
library(readr)
library(dplyr)
#will data frame for raw year and gross
will.gross<-will$movies.50$millions
will.year<-will$movies.50$year
will.raw<-data.frame(will.year, will.gross)
colnames(will.raw) = c("year", "dollar");
clean.will.raw<-na.omit(will.raw)

#denzel data frame for raw year and gross
denzel.gross<-denzel$movies.50$millions
denzel.year<-denzel$movies.50$year
denzel.raw<-data.frame(denzel.year, denzel.gross)
colnames(denzel.raw) = c("year", "dollar");
clean.denzel.raw<-na.omit(denzel.raw)


#inflation data into a frame
result  #inflation table
inflation<-data.frame(result)
inflation<-inflation[rev(1:nrow(inflation)), ]   #reversed columns

helperDeflationRate<-function(yi,yf)
{
  val<-yi/yf
  return(val)
}

getDeflationRate<-function(x)
{
  f<-c()
  for (i in 1:nrow(x))
  {
    xi<-x[i, "dollar"]
    xf<-x[i+1, "dollar"]
    ratio <- helperDeflationRate(xi,xf)
    f<-append(f, ratio)
  }
  f
}
adjusted<-getDeflationRate(inflation)
adjusted<-data.frame(adjusted)     #converted to dataframe
inflation<-cbind(inflation, adjusted)  #added adjusted rate to inflation dataframe

#shift adjust down by 1
inflation<-inflation %>% mutate_at(c("adjusted"), list(lag),n=1)
inflation<-inflation %>% replace(is.na(.),1)
inflation<-inflation[rev(1:nrow(inflation)), ] #reverse order back to ascending

m<-as.numeric(match(clean.will.raw$year, inflation$year))   #matches the will  year to inflation
d<-as.numeric(match(clean.denzel.raw$year, inflation$year)) #matches denzel year to inflation table

getRate2020<-function(x)   #use m to find the percent increase from that m position in inflation
{                #to end of inflation table ... this works!
  final<-c()
  
  for (i in x) 
  {
    adjust<-cumprod(inflation$adjusted[i:101])
    last<-last(adjust)
    final<-append(final, last)
  }
  final
}

final.rate<-sapply(m, FUN=B)  #applies function B to m (will)
final.rate<-sapply(d, FUN=B)  #applies function B to d (denzel)

#add final.rate to will.raw.clean
will.2<-cbind(clean.will.raw,final.rate)
#change will dollar amount with final rate
will.2$dollar_adjusted=will.2$dollar*will.2$final.rate

#final rate to denzel.raw.clean
denzel.2<-cbind(clean.denzel.raw, final.rate)
denzel.2$dollar_adjusted=denzel.2$dollar*denzel.2$final.rate

#testing my code with inflation.txt
test_dat<-read_delim("inflation.txt", delim='|')
n<-as.numeric(match(test_dat$year, inflation$year))  

f.rate<-B(n)
#f.rate<-sapply(n, FUN=B)
f.rate<-data.frame(f.rate)
test_dat2<-cbind(test_dat, f.rate)
test_dat2$testdollar=test_dat2$dollar*test_dat2$f.rate


------------------------------------------------------------
  
## The following code will grab inflation numbers from website and put it into a table form. 
## Teacher provided the txt file of the resulting 101 years
  # data from 1920 to 2020 ... 101 years ...
  infl = "https://www.officialdata.org/us/inflation/2000?endYear=1920&amount=1000000";
# bigger dollar gives a more accurate percent ...

# read the values in "year"/"dollars" using rvest ...
library(rvest);	

infl.html = read_html(infl);

infl.table = infl.html %>%
  html_node(".expand-table-parent") %>%
  html_node(".table-striped") %>%
  html_node("tbody") %>%
  html_nodes("tr");

result = data.frame( matrix(nrow=length(infl.table), ncol=3));
colnames(result) = c("year","dollar","inflation");

for(i in 1:length(infl.table) )
{
  infl.row = infl.table[i]	%>% 
    html_nodes("td") %>%
    html_text();
  
  year = as.numeric(infl.row[1]);
  temp = gsub('$','',infl.row[2],fixed=T);
  temp = gsub(',','',temp,fixed=T);
  dollar = as.numeric(temp);
  temp = gsub('%','',infl.row[3],fixed=T);
  inflation = as.numeric(temp);	
  
  result$year[i] = year;
  result$dollar[i] = dollar;
  result$inflation[i] = inflation;
  
}
result;	
```
